---
title: "p8105_hw2_nt2607"
author: "Niharika Telrandhe"
output: github_document 
date: "2022-10-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## Problem 1
**This problem focuses on NYC Transit data.**

```{r Q1}
nyctransit_df = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(line, starts_with('station'),starts_with('route'), entry, vending, entrance_type, ada, -station_location) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
summary(nyctransit_df)
```
## Short paragraph about the dataset

This dataset contains the following variables : `r names(nyctransit_df)`

My data cleaning steps involved importing the dataset, cleaning of variable names, selecting particular variables and converting the entry variable from character to logical.

The dimensions of the resulting dataset are `r dim(nyctransit_df)`

Yes, these data are tidy.

## Answer the following questions using this data

*How many distinct stations are there?*
There are `r nrow(distinct(nyctransit_df, line, station_name))` distinct stations.
```{r}
nrow(distinct(nyctransit_df, line, station_name))
```

*How many stations are ADA compliant?*
There are `r nrow(filter(nyctransit_df, ada == TRUE))` stations that are ADA compliant.

```{r}
nrow(filter(nyctransit_df, ada == TRUE))
```

*What proportion of station entrances / exits without vending allow entrance?*
```{r}
without_vending = nrow(filter(nyctransit_df, vending == 'NO'))
without_vending

entry_no_vending = nrow(filter(nyctransit_df, vending == 'NO', entry == 'TRUE'))
entry_no_vending

Proportion = entry_no_vending/without_vending
Proportion

```

*Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?*

```{r}
nyctransit_df %>%
  pivot_longer(
    route1:route11,
    names_to = "route_names",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

nyctransit_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_names",
    values_to = "route") %>% 
   filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

## Problem 2
**Read and clean the Mr. Trash Wheel sheet:**

```{r}
mr_trash_df = read_excel("./data/Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls),
         sports_balls = as.integer(sports_balls))
  

```

**Read and clean the Professor Trash Wheel Sheet**

```{r}
Proftrash_df = read_excel("./data/Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") %>%
  janitor::clean_names() %>% 
  drop_na(dumpster)

```
**Combine MR.Trash and Prof Trash**

```{r}
mr_trash_df$dumpster = as.numeric(mr_trash_df$dumpster)
mr_trash_df$year = as.numeric(mr_trash_df$year)
```

**Adding additional variables to both**

```{r}
mr_trash_df$dataset = "MrT"
Proftrash_df$dataset = "ProfT"

join_df = full_join(mr_trash_df, Proftrash_df)

```

**Description**
We combined 'Mr. Trash Wheel' and 'Professor Trash Wheel' to make a combined dataset. 

This has `r nrow(join_df)` rows and `r ncol(join_df)` columns.

It includes 'weight_tons' , 'date'.

The total weight of trash collected by Professor Trash Wheel was `r Proftrash_df %>% pull(weight_tons) %>% sum` 

The total number of sports balls collected by Mr. Trash Wheel in 2020 were `r mr_trash_df %>% filter(year == 2020) %>% pull(sports_balls) %>% sum`


## Problem 3
**First, clean the data in pols-month.csv.**

```{r}
pols_month_df = read.csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(.data = ., month = as.integer(month),
          month = month.name[month],
          year = as.integer(year),
         day = as.integer(day),
         president = case_when(prez_dem =="1" ~ "dem",
                               prez_gop == "1" ~ "gop"
          )) %>% 
  select(-day, -prez_dem, -prez_gop)
head(pols_month_df)
```

Note: The variable prez_gop should only accept 0 or 1 but i can see a few 2's as well which may cause errors.

**Cleaning the data in snp.csv**

```{r}
snp_df = read.csv("./data/fivethirtyeight_datasets/snp.csv") %>%
janitor::clean_names() %>%
  mutate(date = lubridate::mdy(date)) %>%
  separate(date, into = c("year","month","day"), convert = TRUE) %>%
  mutate(month = as.integer(month),
         month = month.name[month],
         year = as.integer(year),
         day = as.integer(day)) %>% 
  select(-day) %>% 
  arrange(year, month) %>% 
  relocate(year, month, close)
head(snp_df)
```

**Cleaning the data in unemployment.csv**

```{r}
unemployment_df = read.csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer( jan:dec,
                names_to = "month",
                values_to = "unemployment_rate") %>% 
  mutate(year = as.numeric(year),
         month = recode(month,
                        jan = "January",
                        feb = "February",
                        mar = "March",
                        apr = "April",
                        may = "May",
                        jun = "June",
                        jul = "July",
                        aug = "August",
                        sep = "September",
                        oct = "October",
                        nov = "November",
                        dec = "December"))
head(unemployment_df)
```

**Merging**

```{r}
join = left_join(pols_month_df, snp_df, by = c("year","month")) %>%
  left_join(unemployment_df, by = c("year", "month"))
```
**Description**

In the `pols_month_df` dataset we have `r nrow(pols_month_df)` rows and `r ncol(pols_month_df)` columns. It contains `r colnames(pols_month_df)` 

In the `snp_df` dataset we have `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. It contains `r colnames(snp_df)`

In the `unemployment_df` dataset we have `r nrow(unemployment_df)` rows and `r ncol(unemployment_df)` columns. It contains `r colnames(unemployment_df)`

The resulting dataset has `r nrow(join)` rows, `r ncol(join)` columns and `r dim(join)` dimensions. the range of years is `r min(join$year)` to `r max(join$year)`. Few of the variables are 'year' 'month' 'gov_gop' 'sen_gop' 'rep_gop' 'gov_dem' 'sen_dem' 'rep_dem' 'president' 'close unemployment_rate'


