---
title: "p8105_hw2_nt2607"
author: "Niharika Telrandhe"
output: github_document 
date: "2022-10-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## Problem 1
This problem focuses on NYC Transit data.

```{r Q1}
nyctransit_df = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(line, starts_with('station'),starts_with('route'), entry, vending, entrance_type, ada, -station_location) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
summary(nyctransit_df)
```
## Short paragraph about the dataset

This dataset contains the following variablea : `r names(nyctransit_df)`

My data cleaning steps involved importing the dataset, cleaning of variable names, selecting particular variables and converting the entry variable from character to logical.

The dimensions of the resulting dataset are `r dim(nyctransit_df)`








